---
title: "ERA5CleanData"
author: "Anna Talucci"
date: "2024-02-13"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview

This script takes extracted ERA5-Land data by year from Google Earth Engine for each raw measurement then reorganizes to for temperature and day of year (DOY) to estimate ALT.

# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
library(data.table)
```



# Data

## vector of data files
```{r}
f <- list.files(path = "../data/ERA5DataEE20250128/",
                pattern = "*.csv", full.names = TRUE)
```

```{r}
f
```

## Individual files
```{r}
f2001 <- read.csv(f[1], header = TRUE) 
f2005 <- read.csv(f[2], header = TRUE) 
f2006 <- read.csv(f[3], header = TRUE) 
f2007 <- read.csv(f[4], header = TRUE) 
f2008 <- read.csv(f[5], header = TRUE) 
f2009 <- read.csv(f[6], header = TRUE) 
f2010 <- read.csv(f[7], header = TRUE) 
f2011 <- read.csv(f[8], header = TRUE) 
f2012 <- read.csv(f[9], header = TRUE) 
f2013 <- read.csv(f[10], header = TRUE) 
f2014 <- read.csv(f[11], header = TRUE) 
f2015 <- read.csv(f[12], header = TRUE) 
f2016CA <- read.csv(f[13], header = TRUE) 
f2016US <- read.csv(f[14], header = TRUE) 
f2017 <- read.csv(f[15], header = TRUE) 
f2018 <- read.csv(f[16], header = TRUE) 
f2019 <- read.csv(f[17], header = TRUE) 
f2020 <- read.csv(f[18], header = TRUE) 
f2023 <- read.csv(f[19], header = TRUE) 
```

```{r}
f2016 = bind_rows(f2016CA, f2016US)
```

## View DF
```{r}
f2001
```
```{r}
names(f2001)
```

## Process

1. pivot longer so column of temperature and date
2. Group by site_id
3. Select dates after day of thaw measurement
4. Select only rows with temp above 0 (zero)
5.  apply calculation

# Functions to clean 

```{r}
selectColDf = function(x)  {
  x %>%
    dplyr::select(plotId, siteId, cntryId, lastNm, submitNm, lat, lon, year, month, day, biome, distur, fireYr, fireId, paired, gtProbe, hitRock, orgDpth, srfH2O, topoPos, slope, vegCvr, msrDoy, msrType, msrDepth,   contains("temperature_2m")) %>%
    mutate(id = row_number()) -> x
  return(x)
}
```

```{r}
cleanDf = function(x)  {
  x %>%
  pivot_longer(
    cols = starts_with("X"),
    names_to = "DateTemp",
    values_to = "TempC",
    values_drop_na = TRUE) %>% 
  separate(DateTemp, c("tempDate", NA)) %>% 
    mutate(plotId = as.character(plotId)) %>%
    mutate(across(tempDate, ~ . |> str_extract("\\d+") |> as.character())) %>%
    mutate(tempDate=ymd(tempDate)) %>%
    mutate(TempDoy = yday(tempDate)) -> x
  return(x)
}
```


##  Apply functions to to each year

### Select Column Function

```{r}
df2001 = f2001 %>% selectColDf()
df2005 = f2005 %>% selectColDf()
df2006 = f2006 %>% selectColDf()
df2007 = f2007 %>% selectColDf()
df2008 = f2008 %>% selectColDf()
df2009 = f2009 %>% selectColDf()
```

```{r}
df2010 = f2010 %>% selectColDf()
df2011 = f2011 %>% selectColDf()
df2012 = f2012 %>% selectColDf()
df2013 = f2013 %>% selectColDf()
df2014 = f2014 %>% selectColDf()
```

```{r}
df2015 = f2015 %>% selectColDf()
df2016 = f2016 %>% selectColDf()
df2017 = f2017 %>% selectColDf()
df2018 = f2018 %>% selectColDf()
df2019 = f2019 %>% selectColDf()
```

```{r}
unique(df2019$submitNm)
```

```{r}
df2020 = f2020 %>% selectColDf()
df2023 = f2023 %>% selectColDf()
```


### Clean function

Note: the warning is referencing parts of the name being discarded in the separate process. This can be ignored for these chunks of code.

```{r}
( clean2001 = df2001 %>% cleanDf() )
```

```{r}
clean2005 = df2005 %>% cleanDf()
clean2006 = df2006 %>% cleanDf()
clean2007 = df2007 %>% cleanDf()
clean2008 = df2008 %>% cleanDf()
clean2009 = df2009 %>% cleanDf()
```

```{r}
clean2010 = df2010 %>% cleanDf()
clean2011 = df2011 %>% cleanDf()
clean2012 = df2012 %>% cleanDf()
clean2013 = df2013 %>% cleanDf()
clean2014 = df2014 %>% cleanDf()
```

```{r}
clean2015 = df2015 %>% cleanDf()
clean2016 = df2016 %>% cleanDf()
clean2017 = df2017 %>% cleanDf()
clean2018 = df2018 %>% cleanDf()
clean2019 = df2019 %>% cleanDf()
```

```{r}
clean2019
```

```{r}
clean2020 = df2020 %>% cleanDf()
clean2023 = df2023 %>% cleanDf()
```

# Function to Identify ALT date for estimating

```{r}
idAltDateFun = function(x)  {
  x %>%
  arrange(id, tempDate) %>%
  group_by(id) %>%
  mutate(RollMean = zoo::rollmean(TempC, k = 5, fill = NA)) %>% 
    filter(TempDoy>=200)  %>%
    ungroup() %>%
    filter(RollMean <=0) %>%
    group_by(lastNm, plotId, siteId, year) %>% # regroup by last name, plot id, and site id
  #filter(site_id== "DFTB") %>%
    arrange(TempDoy) %>% # sort by date of temp measures
  mutate(TempDoy_diff = TempDoy - lag(TempDoy)) %>%
  filter(TempDoy_diff >= 1) %>%
  summarize(minDate = min(TempDoy),
            meanDate = mean(TempDoy),
            maxDate = max(TempDoy)) %>%
    ungroup() -> x
  return(x)
}
```

## Identify ALT date

```{r}
( rm2001 = clean2001 %>% idAltDateFun() )
```

```{r}
rm2005 = clean2005 %>% idAltDateFun()
rm2006 = clean2006 %>% idAltDateFun()
rm2007 = clean2007 %>% idAltDateFun()
rm2008 = clean2008 %>% idAltDateFun()
rm2009 = clean2009 %>% idAltDateFun()
```

```{r}
rm2010 = clean2010 %>% idAltDateFun()
rm2011 = clean2011 %>% idAltDateFun()
rm2012 = clean2012 %>% idAltDateFun()
rm2013 = clean2013 %>% idAltDateFun()
rm2014 = clean2014 %>% idAltDateFun()
```

```{r}
rm2015 = clean2015 %>% idAltDateFun()
rm2016 = clean2016 %>% idAltDateFun()
rm2017 = clean2017 %>% idAltDateFun()
rm2018 = clean2018 %>% idAltDateFun()
rm2019 = clean2019 %>% idAltDateFun()
```

```{r}
rm2019
unique(rm2019$lastNm)
```

```{r}
rm2020 = clean2020 %>% idAltDateFun()
rm2023 = clean2023 %>% idAltDateFun()
```

## Combine Df together

```{r}
combine2001 = merge(clean2001, rm2001[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2005 = merge(clean2005, rm2005[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2006 = merge(clean2006, rm2006[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2007 = merge(clean2007, rm2007[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2008 = merge(clean2008, rm2008[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2009 = merge(clean2009, rm2009[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2010 = merge(clean2010, rm2010[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2011 = merge(clean2011, rm2011[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2012 = merge(clean2012, rm2012[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2013 = merge(clean2013, rm2013[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2014 = merge(clean2014, rm2014[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2015 = merge(clean2015, rm2015[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2016 = merge(clean2016, rm2016[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2017 = merge(clean2017, rm2017[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2018 = merge(clean2018, rm2018[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2019 = merge(clean2019, rm2019[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2020 = merge(clean2020, rm2020[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
combine2023 = merge(clean2023, rm2023[, c('plotId', 'siteId', 'year', 'minDate')], by = c('plotId', 'siteId', 'year'), all.x=TRUE) 
```

```{r}
combine2001
```

```{r}
unique(combine2019$lastNm)
```
```{r}
head(combine2001)
head(combine2005)
head(combine2006)
head(combine2007)
head(combine2008)
head(combine2009)
head(combine2010)
head(combine2011)
head(combine2012)
head(combine2013)
head(combine2014)
head(combine2015)
head(combine2016)
head(combine2017)
head(combine2018)
head(combine2019)
head(combine2020)
head(combine2023)

```
# Bind clean 

Combine yearly data frames; rename columns that were truncated by EE. 

```{r}
( 
combineAll = dplyr::bind_rows(combine2001, combine2005, combine2006, combine2007, combine2008, combine2009, combine2010, combine2011, combine2012, combine2013, combine2014, combine2015, combine2016, combine2017, combine2018, combine2019, combine2020, combine2023) %>%
  mutate(estDoy = minDate-14) %>%
  rename(orgDepth = orgDpth)
)
```


```{r}
names(combineAll)
```

```{r}
unique(combineAll$lastNm)
```


# write to csv

```{r}
write.csv(combineAll,file = "../outputs/era5Clean/RawDataWithERA5.csv", row.names = FALSE)
```



