---
title: "Estimate ALT"
author: "Anna Talucci"
date: "2024-02-13"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```

# Overview

This script estimates end of season Active Layer thickness (ALT) based on time of field measurement, depth of field measurement and ERA5-Land temperature date. 

# Packages

```{r}
library(tidyverse)
library(lubridate)
library(sf)
```

# Data

```{r}
data = read_csv("../outputs/era5Clean/RawDataWithERA5.csv")
```

```{r}
glimpse(data)
```

## Double checks

```{r}
( manies = data %>% filter(lastNm=="Manies") %>%filter(id <=41) )
unique(manies$msrDoy)
```

```{r}
( dieleman = data %>% filter(lastNm=="Dieleman_Baltzer_Turetsky") ) 
```

# Functions to standarize

Formula Parts 
A. Take the square root of the Sum of the total degrees above zero for dates prior to thaw depth measure
B. Take the square root of the Sum of the total degrees above zero for dates prior to active layer thickness date ()

C. Divide B by A
F. Multiply thaw depth measure by C

```{r}
calc_A = function(x) {
  x %>% 
    group_by(year, id, lastNm, msrDoy) %>%
    filter(TempDoy < msrDoy) %>%
  filter(TempC > 0) %>%
    arrange(TempDoy) %>%
  summarize( 
  A_sum = sum(TempC),
  A = sqrt(A_sum)) -> x
  return(x)
}

```


```{r}

calc_B = function(x) {
  x %>% 
    group_by(year, id, lastNm, msrDoy) %>%
    filter(TempDoy < estDoy) %>%
  filter(TempC > 0) %>%
    arrange(TempDoy) %>%
  summarize( 
  B_sum = sum(TempC),
  B = sqrt(B_sum)) -> x
  return(x)
 }
```

### Calculate A & B

```{r}
( calc_A_data = calc_A(data) )
```

```{r}
( calc_B_data = calc_B(data) )
```

### Filter original data 

```{r}
( orgData = data %>% 
  group_by(year, id, lastNm) %>%
  filter(TempDoy == estDoy)
)
  
```

### Join Calculations with Orginial Data

```{r}
( dataCombine = orgData %>% 
    full_join(y=calc_A_data, by=c("year", "id", 'lastNm', 'msrDoy')) %>%
    full_join(y=calc_B_data, by=c("year", "id", 'lastNm', 'msrDoy')) %>%
    mutate(A_sum = ifelse(is.na(A_sum), 0, A_sum),
           A =ifelse(is.na(A), 0, A)) %>%
    mutate(C = B/A) %>%
    mutate(estDepth = round((msrDepth*C), 0)) 
    )
```

```{r}
dataCombine %>% filter(siteId ==("lutose_2000")) %>% filter(plotId ==("4-8")) %>% filter(fireYr==2000) %>% filter(year==2015) %>% write_csv(., "..//outputs/Table3.csv")
```

# Remove outliers, na, and INF

```{r}
dataCombine %>% filter_at(vars(estDepth), all_vars(is.infinite(.)))
dataCombine %>% filter(if_any(estDepth, is.infinite))
dataCombine %>% filter(if_any(estDepth, is.na))
```

# Remove Infinite values

```{r}
( est_infinite = dataCombine %>% 
    filter_at(vars(estDepth), all_vars(is.infinite(.))) %>%
    group_by(submitNm) %>%
    summarize(n= n())
)
```
```{r}
( cleanDataCombine = dataCombine %>% 
    filter_at(vars(estDepth), all_vars(!is.infinite(.))) 
)
```

```{r}
names(cleanDataCombine)
```

```{r}
unique(cleanDataCombine$lastNm)
```

### Write to csv

```{r}
write.csv(cleanDataCombine, file = "../outputs/dataset/FireAltEstimated.csv", row.names=FALSE)
```

