---
title: "Site level aggregate"
author: "Anna Talucci"
date: "2025-01-14"
output: html_document
---

# clear environment
```{r}
rm(list=ls())
```

# Overview

This script aggregates raw measurements to the plot level. 

# Packages

```{r}
library(tidyverse)
library(lubridate)
```


# Data

## Estimated measurements

```{r}
firealt = read_csv("../outputs/dataset/FireAltEstimated.csv")
```

```{r}
firealt
```

# View variables

```{r}
unique(sort(firealt$paired))
```

```{r}
unique(sort(firealt$fireYr))
```
```{r}
unique(sort(firealt$plotId))
```

```{r}
unique(sort(firealt$siteId))
```


```{r}
firealt %>% group_by(submitNm, siteId, plotId, fireYr, year) %>% summarize(rawMeasurements=n())
```

```{r}
firealt %>% group_by(submitNm, siteId, fireYr, year) %>% summarize(rawMeasurments=n())
```
388 sites 9,446 plots by year unique based on location and year sampled.

```{r}
firealt %>% group_by(siteId, fireYr) %>% summarize(n=n())
```

# Summarize raw measurments by submiter with n at plot and site and total raw measurments

```{r}
(
summarizePlotSiteRaw = firealt %>% 
  group_by(submitNm) %>% 
  summarize(numberSites=length(unique(siteId)),       
            numberPlots = length(unique(plotId)),       
            numberMeasurements = n()) 
)
```
```{r}
summarizePlotSiteRaw %>% 
  summarise(siteTotal = sum(numberSites, na.rm = TRUE),
            plotTotal = sum(numberPlots, na.rm = TRUE),
            measurementTotal = sum(numberMeasurements, na.rm = TRUE))
```

# Add time since fire

```{r}
( firealt_tsf = firealt %>% 
  mutate(tsf = year-fireYr) %>%
    mutate(tsfClass = ifelse(tsf %in% 0:3, "0-3",
                                     ifelse(tsf %in% 4:10, "4-10",
                                            ifelse(tsf %in% 11:20, "11-20",
                                                   ifelse(tsf %in% 21:40, "21-40", ">40")))))
)

```

```{r}
firealt_tsf$tsf[firealt_tsf$distur == "unburned"] <- "u"
firealt_tsf$tsfClass[firealt_tsf$distur == "unburned"] <- "unburned"
```

```{r}
sort(unique(firealt_tsf$tsf))
sort(unique(firealt_tsf$tsfClass))
```
```{r}
firealt_tsf$tsf = factor(firealt_tsf$tsf, levels = c("u", "0", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17","19", "20", "21", "29", "30", "31", "32",  "33", "35", "36", "39", "40", "41", "45", "46", "47", "48", "49", "50", "52", "114"))
```

```{r}
firealt_tsf$tsfClass = factor(firealt_tsf$tsfClass, levels = c( "unburned", "0-3", "4-10", "11-20", "21-40", ">40"))
```

# aggregate burn and unburned pairs with summary stats

```{r}
firealt_tsf
```

# Final Plot aggregation

```{r}
( PlotLevel = firealt_tsf %>% 
  group_by(submitNm, distur, siteId, plotId, fireYr, year) %>%
  summarise(
    cntryId = paste(unique(cntryId), collapse = ', '),
    lastNm = paste(unique(lastNm), collapse = ', '),
    lat = mean(lat),
    lon = mean(lon),
    biome = paste(unique(biome), collapse = ', '),
    fireId = paste(unique(fireId), collapse = ', '),
    paired = paste(unique(paired), collapse = ', '),
    gtProde = paste(unique(gtProbe), collapse = ', '),
    hitRock = paste(unique(hitRock), collapse = ', '),
    orgDepth = mean(orgDepth),
    srfH2O = paste(unique(srfH2O), collapse = ', '),
    topoPos = paste(unique(topoPos), collapse = ', '),
    slope = mean(slope),
    vegCvr = paste(unique(vegCvr), collapse = ', '),
    tsf = paste(unique(tsf), collapse = ', '),
    tsfClass = paste(unique(tsfClass), collapse = ', '),
    estDoy = paste(unique(estDoy), collapse = ', '),
    estDepth = mean(estDepth),
    n = n()
    
    )
)
```

```{r}
PlotLevel %>% 
  dplyr::select(-lastNm) %>% 
  filter_all(any_vars(grepl(',', .)))
```
```{r}
names(PlotLevel)
```


```{r}
write.csv(PlotLevel, file = "../outputs/dataset/FireAltEstimatedPlotLevel.csv", row.names=FALSE)
```

