---
title: "Dataset spatial adds"
author: "Anna Talucci"
date: "2024-01-31"
output: html_document
---

```{r clear environment, include=FALSE}
rm(list=ls())
```


# Overview

Add spatial data at the plot level and join plot level spatial adds to raw data. Spatial adds include ecozone and biome from [Resolve Ecoregions](https://hub.arcgis.com/datasets/37ea320eebb647c6838c23f72abae5ef/explore) and permafrost extent.

# Packages

```{r}
library(tidyverse)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(RColorBrewer)
library(cowplot)
library(ggpubr)
library(ggnewscale)
```

# Projection
[Some projection info]https://nsidc.org/data/user-resources/help-center/guide-nsidcs-polar-stereographic-projection)

```{r}
polarProj = "+proj=stere +lat_0=90 +lat_ts=70 +lon_0=-45 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +type=crs"
```

# Data

## Dataset
```{r}
raw = read_csv("../outputs/dataset/FireAltEstimated.csv")
```

```{r}
plots = read_csv("../outputs/dataset/FireAltEstimatedPlotLevel.csv")
```

```{r}
plots
```

## Spatial Data

### Resolve Ecozones

```{r}
resolve = st_read("../data/spatialData/RESOLVE_Ecoregions_and_Biomes/Biomes_and_Ecoregions_2017.shp", "Biomes_and_Ecoregions_2017")
```

```{r}
resolve
```

### Permafrost

```{r}
permafrost = st_read("../data/spatialData/permafrost/permaice.shp", "permaice")
```

```{r}
permafrost
```

# Spatial processing

## Create Spatial points

```{r}
plots
```

### Function to make points
```{r}
df_to_sf <- function(x){
  st_as_sf(x, coords = c("lon","lat"), crs = 4326, remove = FALSE)
}
```

```{r}
( plotsPts = plots %>% 
    df_to_sf() )
```


## Spatial Joins 

### Select Resolve ecozones

```{r}
( resTunBor = resolve %>% 
    filter(REALM %in% c("Palearctic", "Nearctic")) %>% 
    dplyr::select(OBJECTID, ECO_NAME, BIOME_NUM, BIOME_NAME, REALM, ECO_ID, geometry) %>% 
    filter(BIOME_NUM %in% c(11, 6)) %>%
    rename(resId = OBJECTID, resName=ECO_NAME, resBiomNm = BIOME_NUM, resBiome = BIOME_NAME, resRealm = REALM,  resEcoId = ECO_ID)
)
```

### Join pts to ecozones

```{r}
resPtplots = st_join(plotsPts, resTunBor) 
```


```{r}
resPtplots
```

### Reproject to permafrost projection

```{r}
resPtplots_prj = st_transform(resPtplots, crs=st_crs(permafrost))
```



```{r}
resPtplots_prj
```

### Join points to permafrost

```{r}
permResPtplots = st_join(resPtplots_prj, permafrost) 
```

```{r}
permResPtplots
```


### Rename Columns

```{r}
( permResPtPlotsRename = permResPtplots %>% 
  rename(permaExtent= EXTENT) %>%
    select(submitNm:n, resName, resBiome, permaExtent, geometry))
```



### Reproject to WGS

```{r}
permResPtPlots_Wgs = st_transform(permResPtPlotsRename, crs=4326)
```



```{r}
permResPtPlots_Wgs
```

# Data checks

```{r}
unique(permResPtPlots_Wgs$submitNm)
```
```{r}
sort(unique(permResPtPlots_Wgs$resName))
```



```{r}
permResPtPlots_Wgs %>% filter(resName == "Arctic coastal tundra")
permResPtPlots_Wgs %>% filter(siteId == "Jones Old Burn") # checking for mismatched ecozone by pair
permResPtPlots_Wgs %>% filter(submitNm == "Diaz") # checking for mismatched biome by pair/plot observers designation
permResPtPlots_Wgs %>% filter(submitNm == "O'Donnell") # checking for mismatched biome by pair/plot observers designation
permResPtPlots_Wgs %>% filter(submitNm == "Natali") # checking for mismatched biome by pair/plot observers designation
```



## Reassignment based on field observations and plot mismatches

Reassignments are based on discrepancies between field observations a spatial joins. All reassignments were examined on a map. All reassignments were adjacent to polygon boundaries. Reassignments weight field observations higher than spatial joins. Additionally, reassignments are crucial for paired burned/unburned aggregates.

### WOB FireId

```{r}
permResPtPlots_Wgs$resName[permResPtPlots_Wgs$fireId == "WOB"] <- "Arctic foothills tundra"
# convert the 200 pts showing up as Arctic Coastal Tundra to Arctic Foothills Tundra due to pairing; inspected on map to make this decision
```

```{r}
permResPtPlots_Wgs %>% filter(siteId == "Jones Old Burn")
```

### Diaz submitNm

```{r}
permResPtPlots_Wgs %>% filter(submitNm=="Diaz") %>% arrange(siteId)
```

```{r}
permResPtPlots_Wgs$resName[permResPtPlots_Wgs$fireId == "PN6K"] <- "Beringia lowland tundra"
permResPtPlots_Wgs$resBiomeName[permResPtPlots_Wgs$fireId == "PN6K"] <- "Tundra"
permResPtPlots_Wgs %>% filter(submitNm=="Diaz")
```
### O'Donnell SubmitNm

Interior Alaska-Yukon lowland taiga

```{r}
permResPtPlots_Wgs %>% filter(submitNm=="O'Donnell") %>% arrange(siteId)
```

```{r}
permResPtPlots_Wgs$resName[permResPtPlots_Wgs$submitNm == "O'Donnell"] <- "Interior Alaska-Yukon lowland taiga"
permResPtPlots_Wgs$resBiomeName[permResPtPlots_Wgs$submitNm == "O'Donnell"] <- "Boreal Forests/Taiga"
permResPtPlots_Wgs %>% filter(submitNm=="O'Donnell")
```
## Natali

```{r}
permResPtPlots_Wgs %>% filter(submitNm=="Natali") %>% arrange(siteId)
```
Interior Alaska-Yukon lowland taiga
"NCB_NEW" "NCB_OLD" "NCU_NEW" "NCU_OLD"
```{r}
permResPtPlots_Wgs$resName[permResPtPlots_Wgs$siteId == "NCB_NEW"] <- "Interior Alaska-Yukon lowland taiga"
permResPtPlots_Wgs$resBiomeName[permResPtPlots_Wgs$siteId == "NCB_NEW"] <- "Boreal Forests/Taiga"
permResPtPlots_Wgs$resName[permResPtPlots_Wgs$siteId == "NCB_OLD"] <- "Interior Alaska-Yukon lowland taiga"
permResPtPlots_Wgs$resBiomeName[permResPtPlots_Wgs$siteId == "NCB_OLD"] <- "Boreal Forests/Taiga"
permResPtPlots_Wgs$resName[permResPtPlots_Wgs$siteId == "NCU_NEW"] <- "Interior Alaska-Yukon lowland taiga"
permResPtPlots_Wgs$resBiomeName[permResPtPlots_Wgs$siteId == "NCU_NEW"] <- "Boreal Forests/Taiga"
permResPtPlots_Wgs$resName[permResPtPlots_Wgs$siteId == "NCU_OLD"] <- "Interior Alaska-Yukon lowland taiga"
permResPtPlots_Wgs$resBiomeName[permResPtPlots_Wgs$siteId == "NCU_OLD"] <- "Boreal Forests/Taiga"

permResPtPlots_Wgs %>% filter(submitNm=="Natali")
```

### Rechecks

```{r}
sort(unique(permResPtPlots_Wgs$resName))
```
# Dataframe drop geometry

```{r}
plots = permResPtPlots_Wgs %>% st_drop_geometry() %>% select(-resBiomeName)
```

```{r}
names(plots)
```

```{r}
plots
```

# Save files

## write to csv

```{r}
write.csv(plots, file = "../outputs/dataset/2025-02-04_FireAltEstimatedPlotLevel.csv", row.names=FALSE)
```

# Join ecozone and permafrost to raw measurments

```{r}
plots
```
```{r}
( plotsJoin = plots %>% 
    dplyr::select(submitNm, siteId, plotId, distur, fireYr, fireId, year, paired, resName, resBiome, permaExtent, tsf, tsfClass)
)
```

```{r}
( rawData = raw %>% 
    full_join(., plotsJoin, by=c("submitNm", "siteId", "plotId", "distur", "fireYr", "fireId", "year", "paired")) %>%
    dplyr::select(submitNm, lastNm, plotId:cntryId, lat:msrDepth, estDoy, estDepth, tsf, tsfClass, resName:permaExtent ))
```


# Data checks

```{r}
names(rawData)
```

```{r}
unique(rawData$submitNm)
```
```{r}
sort(unique(rawData$resName))
```

# write to csv

```{r}
write.csv(rawData, file = "../outputs/dataset/2025-02-04_FireAltEstimatedRawData.csv", row.names=FALSE)
```